---
title: "Aides Drug"
author: "Andrew Valles Tovar"
date: "Spring 2023"
output: html_document
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
triUmp <-read.csv(file="triUmpsal.csv",header=TRUE)
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?.  If you skimp in this section expect to be marked down in your grade. Paint a picture of the project a couple of paragraphs is expected!-->

A new  preventative aides drug (triUmpsal) is under test in clinical trials. Consider only the DrugUsers, does this drug seem effective?  Hint - use dplyr to obtain that subpopulation with the filter command.


```{r}
druggies <- filter(triUmp,DrugUser=="yes")
```

# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method. Discuss what data manipulations may be needed for example subsetting or other binning procedures and what sort of graphical displays and inferential tools will be used!-->

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming or departing to what is expected by your null hypothesis !-->

```{r}
barchartGC(~infected + treatment,data=druggies)
barchartGC(~infected + treatment,data=druggies, type="percent")

```

Here you have the majority of people who ended up with aids had taken the placebo pill and a significantly smaller amount of people with aids took the drug. On the other column, people who did not have aids had nearly similar results of taking the drug or the placebo.

### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
table2 <- xtabs(~infected + treatment, data=druggies)
rowPerc(table2)
colPerc(table2)
```

For the row data, 24% of people with aids took the drug and 75% of them took the placebo. For those who did not have aids, 47% of the took the drug and 52% took the placebo. For the column data, almost 3% of people who took the drug contracted aids and 97% of them did not get aids. Almost 8% of the people who took the placebo contracted aids and 92% of them did not get aids.

## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
chisq.test(table2)
chisqtestGC(table2)
fisher.test(table2)
```

Based on the inferential data, 34 people who took the drug contracted aids and 106 who took the placebo also contracted aids. A total of 2402 participants were not inflicted with aids. 1143 of them took the drug and 1259 had taken the placebo. There's a 0.0000000423% chance of getting the results indicated by the Chi-square statistic. The Fisher's test states that the odd ratio is 0.35.

# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->
In summary, we can conclude that the small minority of people who had taken the drug, nearly 3%, contracted aids. 8% of the ones who had taken the placebo also contracted aids. A vast majority of particpants who took the drug and placebo did not contract aids, being 97% and 92% of participants respectively. 48% of participants were given the drug and 52% were given the placebo. Based on the data shown, the drug seems to be effective, but there is still a small chance of contracting aids.